% =============================================================================
%  Zhe (Barry) Chen — Curriculum Vitae
%  Compile with XeLaTeX or LuaLaTeX (for fontspec / Unicode support).
%  If you prefer pdfLaTeX, remove the fontspec block and replace ¥ with \textyen
%  (after \usepackage{textcomp}).
% =============================================================================
\documentclass[11pt,a4paper]{article}

% --------- Page geometry ---------
\usepackage[a4paper, top=1in, bottom=1in, left=0.86in, right=0.86in]{geometry}

% --------- Encoding & fonts ---------
\usepackage{fontspec}
% Prefer Palatino Linotype on systems where it is available. macOS commonly
% ships the same family under the shorter Palatino name.
\IfFontExistsTF{Palatino Linotype}
  {\setmainfont{Palatino Linotype}}
  {\IfFontExistsTF{Palatino}{\setmainfont{Palatino}}{}}

% --------- Colour, links, formatting ---------
\usepackage{xcolor}
\definecolor{darkblue}{RGB}{31,73,125}   % Word "Accent 1, darker 25%"
\usepackage[unicode,hidelinks]{hyperref}
\usepackage{bookmark}
\hypersetup{
  colorlinks=true,
  urlcolor=darkblue,
  linkcolor=darkblue,
  bookmarksopen=true,
  bookmarksopenlevel=2,
  bookmarksnumbered=false,
  pdfpagemode=UseOutlines
}

\usepackage{titlesec}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{array}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{ragged2e}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{academicons}
\setlength{\LTpre}{0pt}
\setlength{\LTpost}{0pt}

% --------- Section heading style ---------
\newcommand{\sectionheading}[1]{%
  \begingroup
  \renewcommand{\arraystretch}{1}%
  \setlength{\tabcolsep}{0pt}%
  \setlength{\arrayrulewidth}{1pt}%
  \begin{tabularx}{\linewidth}{@{}X@{}}
  {\fontsize{12}{14}\selectfont\bfseries\color{darkblue}#1}\\[2pt]\hline
  \end{tabularx}%
  \endgroup
}
\titleformat{\section}
  {}
  {}{0pt}{\sectionheading}
\titlespacing*{\section}{0pt}{14pt}{2pt}
\titleformat{\subsection}
  {\normalsize\bfseries\color{darkblue}}
  {}{0pt}{}
\titlespacing*{\subsection}{0pt}{6pt}{4pt}

% --------- PDF outline helpers ---------
\newcounter{cvbookmark}
\newcommand{\cvoutline}[2]{%
  \stepcounter{cvbookmark}%
  \pdfbookmark[#1]{#2}{cvbookmark.\thecvbookmark}%
}
\newcommand{\cvsection}[2]{%
  \cvoutline{1}{#1}%
  \section*{#2}%
}
\newcommand{\cvsubsection}[2]{%
  \cvoutline{2}{#1}%
  \subsection*{#2}%
}

% --------- Helpers ---------
% Highlight self-name in publications (bold + underline)
\newcommand{\me}[1]{\underline{\textbf{#1}}}
\newcommand{\journal}[1]{\textcolor{darkblue}{\textbf{\textit{#1}}}}
\newfontfamily\emailiconfont{Arial Unicode MS}
\newfontfamily\homepageiconfont{Apple Symbols}
\newcommand{\contacticonbox}[1]{\makebox[1.15em][c]{#1}\hspace{0.35em}}
\newcommand{\emailicon}{\contacticonbox{\textcolor{darkblue}{\raisebox{0.02em}{\scalebox{1.32}{{\emailiconfont\char"2709}}}}}}
\newcommand{\homepageicon}{\contacticonbox{\textcolor{darkblue}{\raisebox{-0.03em}{\scalebox{0.9}{{\homepageiconfont\char"1F310}}}}}}
\newcommand{\homepageurl}{zhechen06.github.io}
\newcommand{\scholaricon}{\textcolor{darkblue}{\aiGoogleScholar}}
\newcommand{\scholarstats}{\href{https://scholar.google.com/citations?user=2qgIJwwAAAAJ\&hl=zh-CN}{\scholaricon\hspace{0.25em}\textcolor{darkblue}{Citations}}: 2,089; h-index: 18; hot paper: 1; highly cited paper: 2}

% No paragraph indent, looser line spacing
\setlength{\parindent}{0pt}
\renewcommand{\arraystretch}{1.02}
\hyphenpenalty=10000
\exhyphenpenalty=10000
\emergencystretch=3em
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\fancyfoot[L]{\footnotesize\textcolor{white}{Last updated: May 29, 2026}}
\fancyfoot[R]{\footnotesize Page \thepage\ of \pageref{LastPage}}

% List styling
\setlist{nosep, topsep=1pt, partopsep=0pt, itemsep=0pt, parsep=0pt}
\setlist[enumerate,1]{labelindent=0pt, leftmargin=*, widest=30, labelsep=0.35em, align=right}

% Custom column types
\newcolumntype{D}{>{\RaggedRight\arraybackslash}p{2.75cm}}   % date column
\newcolumntype{Y}{>{\RaggedRight\arraybackslash}p{0.9cm}}    % year column
\newcolumntype{R}{>{\raggedleft\arraybackslash}p{3.5cm}}     % right-aligned role
\newcolumntype{M}{>{\RaggedRight\arraybackslash}X}           % flexible middle

\begin{document}
\linespread{1.18}\selectfont

% =============================================================================
%  HEADER
% =============================================================================
{\fontsize{16}{19}\selectfont\color{darkblue}\textbf{Zhe (Barry) Chen}}\par\vspace{3pt}
{\fontsize{12}{14}\selectfont\textbf{Distinguished Postdoctoral Fellow}}\par
Department of Building Environment and Energy Engineering\par
The Hong Kong Polytechnic University\par
% ZN1001, 181 Chatham Road South, Kowloon, Hong Kong\par
\emailicon\href{mailto:zhechen@polyu.edu.hk}{zhechen@polyu.edu.hk}\par
\homepageicon\href{https://zhechen06.github.io}{\homepageurl}

% =============================================================================
%  WORK EXPERIENCE
% =============================================================================
\cvsection{WORK EXPERIENCE}{WORK EXPERIENCE}
\begin{tabularx}{\textwidth}{@{}D M R@{}}
2025.01--Present &
  \textbf{The Hong Kong Polytechnic University}\newline
  Supervisor: Prof.\ Fu XIAO &
  \textbf{Postdoctoral Fellow} \\
2020.07--2021.07 &
  \textbf{University of Shanghai for Science and Technology}\newline
  Supervisor: Prof.\ Yongbao CHEN &
  \textbf{Research Assistant} \\
\end{tabularx}

% =============================================================================
%  EDUCATION
% =============================================================================
\cvsection{EDUCATION}{EDUCATION}
\begin{tabularx}{\textwidth}{@{}D M R@{}}
2021.09--2024.12 &
  \textbf{The Hong Kong Polytechnic University}\newline
  Supervisor: Prof.\ Fu XIAO &
  \textbf{PhD} \\
2017.09--2020.04 &
  \textbf{Tongji University}\newline
  Supervisor: Prof.\ Peng XU &
  \textbf{MEng} \\
2013.09--2017.06 &
  \textbf{Nanjing University of Aeronautics and Astronautics} &
  \textbf{BEng} \\
\end{tabularx}

% =============================================================================
%  AWARDS & HONORS
% =============================================================================
\cvsection{AWARDS \& HONORS}{AWARDS \& HONORS}
\begin{longtable}{@{}p{0.75cm}@{\hspace{0.20cm}}>{\justifying\arraybackslash}p{\dimexpr\textwidth-0.95cm\relax}@{}}
2025 & \textbf{Champion (Ranking 1/610, AU\$10,000)}, FlexTrack Challenge---Detecting Energy Flexibility in Buildings, AIcrowd, New South Wales Government, and University of Wollongong \\
2025 & \textbf{FCE Awards for Outstanding PhD Theses}, the Hong Kong Polytechnic University (Top 2.5\%) \\
2024 & \textbf{Distinguished Postdoctoral Fellowship}, the Hong Kong Polytechnic University; Finalist, RGC Junior Research Fellow Scheme (the only department nominee) \\
2024 & \textbf{Highly Cited Paper Award}, \textit{Advances in Applied Energy} (IF: 13.8)\newline The most cited paper published in this journal in 2023 \\
2023 & \textbf{Champion (CN¥100,000)}, the First Great Bay Area Data Application Innovation Competition, the Development and Reform Commission of Shenzhen Municipality \\
2023 & \textbf{Excellent Control Award}, the 21st MDV Central Air Conditioning Design and Application Competition, Midea Building Technologies \\
2023 & \textbf{Champion (Ranking 1/331, CN¥20,000)}, the Schneider AI Master Cup Algorithm Competition, Schneider Electric and Baidu PaddlePaddle \\
2023 & \textbf{ESI Hot Paper and Highly Cited Paper}, corresponding author paper titled ``\textit{Physical energy and data-driven models in building energy prediction: A review}'' \\
2023 & \textbf{Champion (Ranking 1/513, CN¥20,000)}, the UNiLAB Smart Energy System Algorithm Competition, UNiLAB and Alibaba Tianchi \\
2022 & \textbf{Paper nominated for the Eni Award 2023}, first author paper titled ``Multi-objective residential load scheduling approach for demand response in smart grid'' \\
2022 & \textbf{Champion}, i-Space: VRAR Contest 2021/22, project titled ``AR-enabled Smart Building Energy Management Platform'', the Hong Kong Polytechnic University \\
2022 & \textbf{Silver Award}, the Global AI Challenge for Building E\&M Facilities, the Electrical and Mechanical Services Department \\
2021 & \textbf{Presidential PhD Fellowship}, the Hong Kong Polytechnic University \\
\end{longtable}

% =============================================================================
%  PUBLICATIONS
% =============================================================================
\cvsection{PUBLICATIONS}{\texorpdfstring{PUBLICATIONS\hfill\normalfont\normalsize\color{black}\scholarstats}{PUBLICATIONS}}

\cvsubsection{Journal Papers}{\color{darkblue}Journal Papers}
\begin{enumerate}
\item \me{Chen Z}, Xiao F*, Guo F. Similarity learning-based fault detection and diagnosis in building HVAC systems with limited labeled data. \journal{Renewable and Sustainable Energy Reviews} 2023;185:113612. (IF: 16.3)

\item \me{Chen Z}, Xiao F*, Guo F, Yan J. Interpretable machine learning for building energy management: A state-of-the-art review. \journal{Advances in Applied Energy} 2023;9:100123. (IF: 13.8, Highly Cited Paper Award)

\item \me{Chen Z}, Chen Y*, He R, Liu J, Gao M, Zhang L. Multi-objective residential load scheduling approach for demand response in smart grid. \journal{Sustainable Cities and Society} 2022;76:103530. (IF: 12.0, Nominated for the Eni Award 2023)

\item \me{Chen Z}, Xiao F*, Chen Y. Multi-objective online optimization of building energy systems for improved control smoothness and efficiency. \journal{Automation in Construction} 2026;181:106604. (IF: 11.5)

\item \me{Chen Z}, Zhang J, Xiao F*, Xu K, Chen Y. Development of a probabilistic cooling load prediction-based robust chiller sequencing strategy and its real-world implementation. \journal{Applied Energy} 2025;382:125213. (IF: 11.0)

\item \me{Chen Z}, Xiao F*, Xiao Z, Chen Y. Bridging the gap between data-driven baselines and energy saving uncertainty for building retrofit. \journal{Energy} 2025;340:139292. (IF: 9.4)

\item \me{Chen Z}, Zhang J, Xiao F*, Xu K, Chen Y. Physically consistent data-driven optimal sequencing strategy for variable speed pumps in large building chiller plants. \journal{Building and Environment} 2025;281:113177. (IF: 7.6)

\item \me{Chen Z}, Chen Y*, Xiao T, Wang H, Hou P. A novel short-term load forecasting framework based on time-series clustering and early classification algorithm. \journal{Energy and Buildings} 2021;251:111375. (IF: 7.1)

\item \me{Chen Z}, Xu P*, Feng F, Qiao Y, Luo W. Data mining algorithm and framework for identifying HVAC control strategies in large commercial buildings. \journal{Building Simulation} 2021;14:63--74. (IF: 5.9)

\item \me{Chen Z}, Zhang J, Xiao F*, Madsen H, Xu K. Probabilistic machine learning for enhanced chiller sequencing: A risk-based control strategy. \journal{Energy and Built Environment} 2025;6:783--95. (ESCI indexed, Scopus CiteScore: 17.3)

\item Chen Y, \me{Chen Z}*. Short-term load forecasting for multiple buildings: A length sensitivity-based approach. \journal{Energy Reports} 2022;8:14274--88. (IF: 5.1, Corresponding author)

\item Chen Y*, Guo M, Chen Z, \me{Chen Z}*, Ji Y. Physical energy and data-driven models in building energy prediction: A review. \journal{Energy Reports} 2022;8:2656--71. (IF: 5.1, Corresponding author, ESI Hot Paper, ESI Highly Cited Paper)

\item Tang H, \me{Chen Z}, Li H, Wang S*. Risk-aware optimal dispatch of resource aggregators integrating NGBoost-based probabilistic renewable forecasting and bi-level building flexibility engagements. \journal{Engineering} 2025;53:76--89.

\item Xu K, \me{Chen Z}, Xiao F*, Zhang J, Zhang H, Ma T. Semantic model-based large-scale deployment of AI-driven building management applications. \journal{Automation in Construction} 2024;165:105579.

\item Chen Y, \me{Chen Z}, Xu P*, Li W, Sha H, Yang Z, et al. Quantification of electricity flexibility in demand response: Office building case study. \journal{Energy} 2019;188:116054.

\item Chen Y, \me{Chen Z}, Ding Y, Du Y*, Li B, Liu J, et al. Towards Transformer Winding Deformation Fault Diagnosis under Data Scarcity: Integrating Frequency Response Analysis with Metric Learning. \journal{IEEE Transactions on Dielectrics and Electrical Insulation} 2026:1--1.

\item Chen Y*, \me{Chen Z}, Yuan X, Su L, Li K*. Optimal control strategies for demand response in buildings under penetration of renewable energy. \journal{Buildings} 2022;12:371.

\item Chen Y, Xu P*, \me{Chen Z}, Wang H, Sha H, Ji Y, et al. Experimental investigation of demand response potential of buildings: Combined passive thermal mass and active storage. \journal{Applied Energy} 2020;280:115956.

\item Song Z, Xiao F*, \me{Chen Z}, Madsen H. Probabilistic ultra-short-term solar photovoltaic power forecasting using natural gradient boosting with attention-enhanced neural networks. \journal{Energy and AI} 2025;20:100496.

\item Chen Y*, Yang Q*, \me{Chen Z}, Yan C, Zeng S, Dai M. Physics-informed neural networks for building thermal modeling and demand response control. \journal{Building and Environment} 2023;234:110149.

\item Chen Y*, Wang H*, \me{Chen Z}. Sensitivity analysis of physical regularization in physics-informed neural networks (PINNs) of building thermal modeling. \journal{Building and Environment} 2025;273:112693.

\item Xiao Z, Zhang J, \me{Chen Z}, Xiao F*. Probabilistic building energy savings verification via triplet network-based adaptive reference day selection. \journal{Energy and Buildings} 2026;364:117664. (IF: 7.1)

\item Chen Y*, Chen Z, \me{Chen Z}, Yuan X. Dynamic modeling of solar-assisted ground source heat pump using Modelica. \journal{Applied Thermal Engineering} 2021;196:117324.

\item Yuan H, Chen Y*, \me{Chen Z}. Investigation on Electricity Flexibility and Demand-Response Strategies for Grid-Interactive Buildings. \journal{Buildings} 2025;15:4368.

\item Xiao T, Xu P*, Ding R, \me{Chen Z}. An interpretable method for identifying mislabeled commercial building based on temporal feature extraction and ensemble classifier. \journal{Sustainable Cities and Society} 2022;78:103635.

\item Chu Y, Xu P*, Li M, \me{Chen Z}, Chen Z, Chen Y, et al. Short-term metropolitan-scale electric load forecasting based on load decomposition and ensemble algorithms. \journal{Energy and Buildings} 2020;225:110343.

\item Xiao Z, Zhang J*, Xiao F*, \me{Chen Z}, Xu K, So PM, et al. An AI-enabled optimal control strategy utilizing dual-horizon load predictions for large building cooling systems and its cloud-based implementation. \journal{Energy and Buildings} 2025;330:115352.

\item Yuan H, Chen Y*, Di Gangi A, \me{Chen Z}. Energy Flexibility Evaluation for Building Passive Thermal Storage Mass. \journal{Energies} 2026;19:1035. (IF: 3.2)

\item Han C, Chen Y*, Wang H, Zhan S, \me{Chen Z}. Development and validation of a data quality assessment framework for machine learning-based building load prediction. \journal{Applied Energy} 2026;415:127925.

\item Sha H, Xu P*, Hu C, Li Z, Chen Y, \me{Chen Z}. A simplified HVAC energy prediction method based on degree-day. \journal{Sustainable Cities and Society} 2019;51:101698.

\item Li A, Zhang J, Xiao F*, Fan C, Yu Y, \me{Chen Z}. Design information-assisted graph neural network for modeling central air conditioning systems. \journal{Advanced Engineering Informatics} 2024;60:102379.

\item Gu J, Xu P*, Pang Z, Chen Y, Ji Y, \me{Chen Z}. Extracting typical occupancy data of different buildings from mobile positioning data. \journal{Energy and Buildings} 2018;180:135--45.
\end{enumerate}

\cvsubsection{Conference Papers and Presentations}{\color{darkblue}Conference Papers and Presentations}
\begin{enumerate}
\item \me{Chen Z}, Xiao F*. A novel demand response detection and capacity prediction framework. \textit{Applied Energy Symposium: AIXEnergy}, May 2026, Ningbo, China.

\item \me{Chen Z}, Xiao F*. Reliably estimating energy savings using data-driven models as the baseline. \textit{International Conference on Building and Simulation} \textbf{(BAS)}, July 2025, Borlänge, Sweden.

\item \me{Chen Z}, Xiao F*. Robust comparison of optimization algorithms for optimal chiller loading. \textit{International Conference on Sustainable Development in Building and Environment} \textbf{(SuDBE)}, Aug.\ 2023, Espoo, Finland.

\item \me{Chen Z}, Xiao F*. Similarity learning-based FDD in building HVAC systems. \textit{The 3rd International Conference for Global Chinese Academia on Energy and Built Environment International} \textbf{(CEBE)}, July 2023, Shanghai, China.

\item \me{Chen Z}, Chen Y*, Mohtaram S. Quantifying HVAC electrical flexibility from building thermal mass: a case study of the DOE reference building. \textit{The 16th ROOMVENT Conference} \textbf{(ROOMVENT)}, Sept.\ 2022, Xi'an, China.

\item \me{Chen Z}, Xiao F*. Interpretable machine learning for AHU fault detection and diagnosis: Rule extraction from black-box models. \textit{The 14th International Conference on Applied Energy} \textbf{(ICAE)}, Aug.\ 2022, Bochum, Germany.

\item \me{Chen Z}, Chen Y*. A multi-objective optimization-based load scheduling approach for demand response in smart grid. \textit{The 2nd International Conference for Global Chinese Academia on Energy and Built Environment International} \textbf{(CEBE)}, July 2021, Chengdu, China.

\item \me{Chen Z}, Xu P*, Chen Y. A peer-to-peer electricity system and its simulation. \textit{The 4th Asia Conference of International Building Performance Simulation Association} \textbf{(ASIM)}, Dec.\ 2018, Hong Kong, China.
\end{enumerate}

\cvsubsection{Conference Session Chairs}{\color{darkblue}Conference Session Chairs}
\begin{enumerate}
\item \textit{The 2nd International Conference on Digital Intelligence for Energy Systems} \textbf{(ICDIES)}, Oct.\ 2026, Wollongong, Australia.
\item \textit{Applied Energy Symposium: AIXEnergy}, May 2026, Ningbo, China.
\end{enumerate}

\cvsubsection{Invited Talks}{\color{darkblue}Invited Talks}
\begin{enumerate}
\item Demand Response Detection and Capacity Prediction using Ensemble Learning. \textit{IEEE International Conference on Energy Technologies for Future Grids}, Dec.\ 2025, Wollongong, Australia. (invited as the winner of FlexTrack Challenge)
\item Embarking on your research journey: My experience at PolyU BEEE. \textit{Research Postgraduate Student Orientation, Department of Building Environment and Energy Engineering, The Hong Kong Polytechnic University}, Sept.\ 2025, Hong Kong, China. (invited as PhD graduate representative)
\end{enumerate}

% =============================================================================
%  PATENTS
% =============================================================================
\cvsection{PATENTS}{PATENTS}
\begin{enumerate}
\item A demand response control method for multi-energy residential buildings, CN202110226687.0.
\item A chiller control method, apparatus, equipment, and storage medium, CN118189539B.
\item An automatic generation method for schematic diagrams of HVAC water systems based on topological layered abstraction, CN202010507609.3.
\item A path optimization method for water systems in air conditioning systems, CN202011126801.4.
\item An optimization algorithm for pipe shafts, CN202011074593.8.
\item An automatic equipment selection algorithm for HVAC systems, CN202110515557.9.
\end{enumerate}

% =============================================================================
%  RESEARCH PROJECTS
% =============================================================================
\cvsection{RESEARCH PROJECTS}{RESEARCH PROJECTS}
\begin{longtable}{@{}p{2.7cm} >{\RaggedRight\arraybackslash}p{\dimexpr\textwidth-2.7cm-2\tabcolsep\relax}@{}}
2025--2027 &
  \textbf{Development of Smart and Scalable Data-driven Management Technologies for Building HVAC Systems towards Carbon Neutrality}\newline
  PolyU Distinguished Postdoctoral Fellowship Scheme, HK\$841{,}040, P0053872\newline
  Role: Fellowship Awardee \\
2022--2024 &
  \textbf{Smart Data-Driven Building Management Framework for Environmental and Sustainability Applications}\newline
  Innovation and Technology Fund of Hong Kong, ITP/002/22LP\newline
  Role: Lead researcher for demonstration of data-driven building management technologies in large buildings \\
2021--2024 &
  \textbf{Development of Smart Energy Management Technologies for Complex Building Energy Systems in High-Density Cities}\newline
  National Key Research and Development Program of China, 2021YFE0107400\newline
  Role: Lead researcher for data-driven modelling and optimization for building energy systems \\
2017--2021 &
  \textbf{Distributed Integrated Energy System Joint Simulation Technologies}\newline
  National Key Research and Development Program of China, 2017YFB0903404\newline
  Role: Lead researcher for distributed integrated energy system modeling \\
\end{longtable}

% =============================================================================
%  ACADEMIC SERVICE
% =============================================================================
\cvsection{ACADEMIC SERVICE}{ACADEMIC SERVICE}
\begin{tabularx}{\textwidth}{@{}p{2.7cm} M@{}}
Editorial Board & Young Editorial Board Member of \textit{Carbon Neutrality} (IF: 12.5, Q1) \\
Guest Editor    & Special issue of \textit{Buildings} (ISSN 2075-5309)\newline
                  \textit{Smart and Sustainable Buildings: Advancing towards Net-Zero and Intelligent Control} \\
Peer Reviewer   & Invited reviewer for journals including: \textit{Applied Energy, Automation in Construction, Energy Conversion and Management, Journal of Cleaner Production, Advanced Engineering Informatics, Engineering Applications of Artificial Intelligence, Computers in Industry, Energy, Building and Environment, Applied Thermal Engineering, Energy and Buildings, Building Simulation, iScience} \\
Membership      & International Building Performance Simulation Association\newline
                  International Society of Energy and Built Environment \\
\end{tabularx}

% =============================================================================
%  TEACHING EXPERIENCE
% =============================================================================
\cvsection{TEACHING EXPERIENCE}{TEACHING EXPERIENCE}
\begin{tabularx}{\textwidth}{@{}p{2.7cm} X r@{}}
Advising & BSE3716 \& BSE4726 (Capstone Research Project), 2025--2026 & 6 undergraduates \\
         & BSE3713 \& BSE4725 (Capstone Research Project), 2023--2024 & 2 undergraduates \\
\end{tabularx}

\end{document}
